Orienting the causal relationship between imprecisely measured traits using genetic instruments

### 2017-03-13 docs/manuscript.tex
updated

\setlength{\droptitle}{-2em}
  \title{Orienting the causal relationship between imprecisely measured traits
using genetic instruments}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{}
  \preauthor{}\postauthor{}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{11 March 2017}

@@

where \(\hat{y} = \hat{\beta}_x x\). MR analysis estimates the causal
influence of \(x\) on \(y\) by using the instrument as a proxy for
\(x\), such that

@@

We demonstrate this new method by evaluating the causal relationships of
458 known associations between DNA methylation and gene expression
levels using summary level data. The causal direction is heavily
influenced by how much measurement error is present in the different
assaying platforms. For example, if DNA methylation measures typically
have higher measuremet error than gene expression measures then our
analysis suggests that DNA methylation levels would be more often the
causal factor in the association. Indeed, previous studies which have
evaluated measurement error in these platforms do support this position
(45,46), though making strong conclusions for this analysis is difficult
because measurement error is likely to be study specific.

>>>>

\setlength{\droptitle}{-2em}
  \title{Orienting the causal relationship between imprecisely measured traits
using genetic instruments}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{}
  \preauthor{}\postauthor{}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{13 March 2017}

@@

where \(\hat{y} = \hat{\beta}_x x\) and assuming no intercept
\(y - \hat{y} = \epsilon_x\). MR analysis estimates the causal influence
of \(x\) on \(y\) by using the instrument as a proxy for \(x\), such
that

@@

We demonstrate this new method by evaluating the causal relationships of
458 known associations between DNA methylation and gene expression
levels using summary level data. The causal direction is heavily
influenced by how much measurement error is present in the different
assaying platforms. For example, if DNA methylation measures typically
have higher measuremet error than gene expression measures then our
analysis suggests that DNA methylation levels would be more often the
causal factor in the association. Indeed, previous studies which have
evaluated measurement error in these platforms do support this position
(45,46), though making strong conclusions for this analysis is difficult
because measurement error is likely to be study specific. We also
haven't accounted for the influence of winner's curse, which can inflate
estimates of the variance explained by SNPs, with higher inflation
expected amongst lower powered studies. Using p-values for genetic
associations from replication studies will mitigate this problem.
